
---
title: "Bootstrapped quantiles"
subtitle: "The use of bootstrapping to estimate the required number of iterations for a given quantile precision when conducting MSE"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  word_document:
  mathjax: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache=TRUE,
  cache.path='cache/qboot/',
  fig.path  ='tex/qboot-',
  fig.align='center', comment=NA,
  message=FALSE, warning=FALSE, echo=TRUE,
  fig.width=6, fig.height=4.5, dpi=300)

iFig=0
```

## Setup
  
```{r}
# packages
library(FLCore)
library(ggplotFL)
library(reshape2)
# functions
source("../R/qboot.R")
source("../R/qbootplot.R")
nboot <- 99
```


# 1. Introduction

MSEs are often used to compare the performance of alternate harvest control rules (HCRs) in terms of variables of interest (e.g. maximum yield, yield variation, and spawning stock biomass). Multiple iterations are conducted in order to take into account various types of variation (e.g. recruitment, observation error, implementation error). An important aspect of HCR testing is in the estimation of the liklihood of particular outcomes occuring.  For example, HCRs frequently used in ICES are based on
this is often centered around mainatining stock levels above a given reference spawning stock biomass (SSB) associated with depleated productivity.

The proposed bootstrapping method estimates errors of distribution quantiles for these variables of interest, which can be used as diagnostic metrics for estimate confidence or in the prediction of the number of iterations required for achieving a desired level of precision. A practical application of the method is to more efficiently identify most promising HCRs, allowing for more economical allocation of computation resources. 


# 2. Methods 

# 2.1. Example dataset


```{r q~niter}
load(file="../data/codMSE_A.Rdata")

SSB <- A["SSB",,, "0.31", "150000"]
ssb <- FLCore::FLQuant(SSB[1,], dim = c(1,dim(SSB)[1],1,1,1,1000), dimnames = dimnames(SSB))
ssb[] <- SSB

L <- FLQuants("iter = 50" = ssb[,,,,,1:50], "iter = 200" = ssb[,,,,,1:200], "iter = 1000" = ssb)
plot(L)
```

**Figure `r iFig=iFig+1; iFig`.** Quantile distributions of spawning stock biomass for North Sea cod  differing number of iterations. Darker shaded areas shows the 25\% and 75\% quantiles, with median indicated by a solid line, and 5\% and 95\% quantiles as dashed lines.




```{r one_hcr}
op <- par(mar = c(2, 3.5,1,1), mgp = c(2,0.5,0), ps = 10)
matplot(x = as.numeric(dimnames(SSB)$year), y = SSB, 
  t = "l", xlab = "", lty = 1, ylab = "SSB [t]", col = rgb(0,0,0,0.1))
par(op)
```



# 2.2. Bootrapping procedure

Bootstrapping approach for quantile distributions (estimate (median), and CIs (e.g. 0.025, 0.975))

Error metric (*median absolute deviation*, MAD):
$$MAD = median(\left| x-median(x) \right|)$$

Relative error
$$RE = \frac{MAD}{median(x)} * 100$$


# 2.3. Predictive model

Prediction of number of iterations required for a given level of relative error (`lm(log(niter) ~ log(rerr) + quant)`).


# 2.4. Application in optimal harvest control rule selection 

3. Use of bootstrapped metrics in algorithm to more efficiently focus MSEs on most promising HCRs.


Algorithm:

1. Define minimum ($i_{min}$), maximum ($i_{max}$) and step size ($i_{step}$) for MSE iterations.
2. Define target relative error ($RE_{targ}$).
3. For each HCR ($HCR_h$), run $i \leftarrow i_{min}$ iterations 
4. Evaluate the quantile of interest with bootstrapping approach and estimate median, confidence intervals, and relative error.
5. If all $RE_h<RE_{targ}$ or $i<i_{max}$, **stop**
6. Perform selection criteria:
+ 6a. Flag $h$ when upper confidence interval of 5\% quantile is greater than Btrigger.
+ 6b. Rank median catches of remaining $h$. 
+ 6c. Flag $h$ that are not in top ranked $h$ (75\% or a minimum of 30).
7. Conduct additional iterations ($i \leftarrow i+i_{step}$) on un-flagged $HCR_h$.
8. Return to step 4. 


# 3. Results





# 3.1. Performance of prediction model


```{r, pred}
# Convert to matrix (year by iteration)
Xfull <- array(ssb, dim = dim(ssb)[c(2,6)], dimnames = dimnames(ssb)[c(2,6)])
Xfull <- t(Xfull) # transpose to make iters rows

X <- array(ssb[1,tail(seq(dim(ssb)[2]),10),1,1,1,], dim = c(10,1000))
dimnames(X) <- list(year = tail(dimnames(ssb)[[2]],10), iter = dimnames(ssb)[[6]]) # add 
X <- t(X) # transpose to make iters rows

# run bootstrapping, model mean
set.seed(1)
resMean <- qboot(X, nboot = nboot, quant = c(0.05, 0.5, 0.95),
  niter = round(exp(seq(log(nrow(X))*0.5, log(nrow(X)), length.out = 10))),
  aggfun = "mean", rerrTarg = 1, verbose = FALSE)
set.seed(1)
resMax <- qboot(X, nboot = nboot, 
  niter = round(exp(seq(log(nrow(X))*0.5, log(nrow(X)), length.out = 10))),
  aggfun = "max", rerrTarg = 1, verbose = FALSE)

op <- par(mfrow = c(1,1), mar=c(3,3,2,0.5), ps=10, mgp = c(2,0.5,0))
qbootplot(resMean, col = pals::brewer.set1(3))
mtext("Mean yearly RE", side = 3, line = 0.25)
# qbootplot(resMax)
# mtext("Maximum yearly RE", side = 3, line = 0.25)
par(op)

```

**Figure `r iFig=iFig+1; iFig`.** Log-log plot of mean (left panel) and max (right panel) bootstrapped relative error (MAD) versus number of iterations by quantile. Predicted linear regressions are shown by solid lines, and predicted number of iterations needed for the target error level are shown by dashed lines, with values in the legend.





```{r ci~niter}
par(mar=c(3,3,2,0.5), ps=10, mgp = c(2,0.5,0))
# make dataframe of subset of ci data
set.seed(1)
resMeanFull <- qboot(Xfull, nboot = nboot, 
  niter = round(exp(seq(log(nrow(Xfull))*0.5, log(nrow(Xfull)), length.out = 10))),
  aggfun = "mean", rerrTarg = 1, verbose = FALSE)

res <- resMeanFull
res2 <- expand.grid(dimnames(res$est))
# res2$niter <- factor(round(as.numeric(as.character(res2$niter))), 
#   levels = ac(sort(as.numeric(ac(levels(res2$niter))))))
res2$niter <- as.numeric(ac(res2$niter))
res2$quant <- as.numeric(ac(res2$quant))


res2$year <- as.numeric(as.character(res2$Var1))
res2$est <- c(res$est)
res2$cilow <- c(res$cilow)
res2$ciup <- c(res$ciup)
res2 <- subset(res2, niter %in% unique(res2$niter)[c(1,5,10)] & 
    quant %in% unique(res2$quant)[c(1,3,5)])
res2$niter <- factor(res2$niter, levels = sort(unique(res2$niter)))
res2$quant <- factor(res2$quant, levels = sort(unique(res2$quant)))

ci <- c(res2$cilow, res2$ciup)
plot(est ~ year, data = res2, ylim = range(ci), t="n", xlab = "",
  ylab = "Spawning stock biomass (SSB)")
rect(xleft = min(tail(unique(res2$year), 10)),
  xright = max(tail(unique(res2$year), 10)), ybottom = 0, ytop = 1e6,
  border = NA, col = "grey90")
COL <- pals::brewer.set1(length(levels(res2$quant)))
for(i in seq(levels(res2$quant))){
  for(j in seq(levels(res2$niter))){
    dat.ij = subset(res2, quant == levels(res2$quant)[i] &
      niter == levels(res2$niter)[j])
    # dat.ij[order(dat.ij$year),]
    polygon(x = c(dat.ij$year, rev(dat.ij$year)), 
      y = c(dat.ij$cilow, rev(dat.ij$ciup)),
      col = adjustcolor(COL[i], 0.3), border = NA)# , lty = j+1)
  }
  lines(est ~ year, subset(res2, 
    quant == levels(res2$quant)[i] & niter == 1000), col=1)
}
abline(h = 107000, lty = 3)
box()
```

**Figure `r iFig=iFig+1; iFig`.** True quantiles (lines; 5, 50, & 95\%) and their confidence intervals (95\%, shaded area) for varying sample sizes (`r paste("n =", paste(levels(res2$niter), collapse=", "))`).



```{r, pred_max~cumiter}
cumiter <- seq(100, 1000, 100) # c(250, 500, 750, 1000, 2000, 5000)
res <- vector("list", length(cumiter))
set.seed(1)
for(i in seq(res)){
  set.seed(1)
  res.i <- qboot(X[seq(cumiter[i]),], nboot = 499, aggfun = "mean", 
    rerrTarg = 1, verbose = FALSE)
  res.i$fitpred$cumiter <- cumiter[i]
  res[[i]] <- res.i$fitpred
}
res <- do.call("rbind", res)

ggplot(subset(res, quant %in% c("0.95", "0.5", "0.05"))) + 
  aes(x = cumiter, y = niter, group = quant, colour = quant) + 
  geom_line() + geom_point() + 
  xlab("Cumulative iterations completed") + 
  ylab("Iterations required for 1% RE") +
  scale_colour_brewer(palette = "Set1", direction = 1,
    aesthetics = "colour")
```

**Figure `r iFig=iFig+1; iFig`.** Influence of cumulative sample size on predicted number of iterations needed for a given error level (1\%).





# 3.2. Performance of optimal harvest control rule selection



# 4. Discussion




The bootstrapping approach allows for two main outputs: 

1. Quantile estimates by year and number of iterations, associated standard error and confidence limits
2. Prediction of the number of iterations required for a given level of relative standard error (standard error / mean * 100).

The first output provides a robust estimation quantile estimates by number of iterations through boostrapping with replacement and provides two measures of the confidence of the estimate. The first is standard error ($se$), assuming a roughly normal distribution of the sample estimates as dictated by the central limit theorem. Thus, confidence intervals could be calculated with $se$ and cumulative distribution function. A second approach to determining confidence intervals is a direct estimation from the estimation distribution. This is also provided in the form of confidence limits.

The latter output is based on the known relationship between *se* and sample size ($n$), $log(n)=\alpha + \beta~log(se)$, where $\beta \approx -\frac{1}{2}$. The `qboot` function uses relative standard error ($rsa$) in fitting this relationship in order to standardize among years of the data set since, in the case of MSE, short-term values may be of a different magnitude than long-term equilibrium values. Thus, using the linear relationship `log(niter)~log(rsa)`, where *niter* are the number of samples used in the bootstrap estimation, the model is able to approximate the number of iterations required in order to achieve a given level of *rsa*.





## A synthetic example



<!-- **Figure `r iFig=iFig+1; iFig`.** Quantile distributions of spawning stock biomass for a synthetic stock by differing number of iterations. Darker shaded areas shows the 25\% and 75\% quantiles, with median indicated by a solid line, and 5\% and 95\% quantiles as dashed lines. -->


## Predicting number of iterations for a given error level

The following shows mean relative standard error across years by defined quantile and number of iterations. This is related to 'Prob1' risk estimation.



<!-- **Figure `r iFig=iFig+1; iFig`.** Log-log plot of mean bootstrapped relative standard error (rse) versus number of iterations by quantile. Predicted linear regressions are shown by solid lines, and predicted number of iterations needed for the target error level are shown by dashed lines, with values in the legend. -->

The following shows max rse across years by defined quantile and number of iterations. This is related to 'Prob3' risk estimation.


<!-- **Figure `r iFig=iFig+1; iFig`.** Log-log plot of max bootstrapped relative error (rse) versus number of iterations by quantile. Predicted linear regressions are shown by solid lines, and predicted number of iterations needed for the target error level are shown by dashed lines, with values in the legend. -->










## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* ggplotFL: `r packageVersion('ggplotFL')`
* reshape2: `r packageVersion('reshape2')`
* **Compiled**: `r format(Sys.Date(), '%Y-%b-%d')`



